---
title: "Coding Assignment 2"
author: "Group 6"
date: "Due: 2024-10-20 23:59"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
#Put any packages you need here
knitr::opts_chunk$set(echo = TRUE)
```

A Florida health insurance company wants to predict annual claims for individual clients. The company pulls a random sample of 100 customers. The owner wishes to charge an actuarially fair premium to ensure a normal rate of return. The owner collects all of their current customer’s health care expenses from the last year and compares them with what is known about each customer’s plan. 

The data on the 100 customers in the sample is as follows:

-	Charges: Total medical expenses for a particular insurance plan (in dollars)
-	Age: Age of the primary beneficiary
-	BMI: Primary beneficiary’s body mass index (kg/m2)
-	Female: Primary beneficiary’s birth sex (0 = Male, 1 = Female)
-	Children: Number of children covered by health insurance plan (includes other dependents as well)
-	Smoker: Indicator if primary beneficiary is a smoker (0 = non-smoker, 1 = smoker)
-	Cities: Dummy variables for each city with the default being Sanford

Answer the following questions using complete sentences and attach all output, plots, etc. within this report.


```{r dataset, include=FALSE}
InsuranceData <- read.csv("C:\\UserData\\z002r7kt\\OneDrive - Siemens AG\\Reference Documents\\Learnings\\Fall-24_ECO6416_Group6\\Insurance_Data_Group6.csv")
```


## Question 1

Generate this regression equation:

$$Charges = \beta_{0}+\beta_{1}*Age+\beta_{2}*BMI+\beta_{3}*Children + \beta_{4}*Female + \beta_{5}*Smoker + \beta_{6}*WinterSprings + \beta_{7}*WinterPark + \beta_{8}*Oviedo$$

```{r q1}
model <- lm(Charges ~ Age + BMI + Children + Female + Smoker + WinterSprings + WinterPark + Oviedo, data = InsuranceData)

summary(model)


```
  
Regression equation written out clearly:  

Charges = -13716 + 282.33(Age) + 382.48(BMI) -152.02(Children) -1187.23(Female) + 27885.82(Smoker) -515.44(WinterSprings) -2003.37(WinterPark) -1111.73(Oviedo)     
  
***  
Which independent variables test significant in a two-tailed test? Use $\alpha = 0.2$

Independent variables Age, BMI, Smoker and WinterPark test significant in a two-tailed test with $\alpha = 0.2$  

***
## Question 2

Interpret the coefficients of all the independent variables that tested significant.

The independent variables Age, BMI, Smoker and WinterPark test significant in a two-tailed test with $\alpha = 0.2$ This means that Age, BMI, Smoker and WinterPark variables have a statistically significant effect on the dependent variable, Charges, while the variables Children, Gender, WinterSprings and Oviedo are not significant. The coefficients for the statistically significant variables show the change in Total medical expenses for a particular insurance plan (in dollars) for every change in their respective 1 unit. For example, the total medical expenses for a particular insurance plan increases by ~$282 for every 1 year older the primary beneficiary is. 

## Question 3

The insurance agent returns with another client:

- Age: 35
- BMI: 26
- Children: 1
- Gender: Female
- Smoker: No
- City: Sanford

Using the regression equation above, predict the amount of medical expenses associated with this policy. (Provide a 95% confidence interval as well)

```{r q3}

Charges <- data.frame(Age=35,
                      BMI=26,
                      Children=1,
                      Female=1,
                      Smoker=0,
                      WinterSprings=0,
                      WinterPark=0,
                      Oviedo=0)

predict(model, newdata = Charges,interval = "prediction")
predict(model, newdata = Charges,interval = "confidence")

```

## Question 4

Compare the model from question 1, to a model that does not include the group of city dummies. Does including the cities reduce the SSE by a statistically significant amount? (check the partial F-test for significance of the dummies as a group). Use $\alpha = 0.05$

```{r q4}

```


